\textbf{In this paper, we aimed to analyse flight delay and cancellation trends using big data technologies such as Hadoop and Spark. By leveraging data from the U.S. Department of Transportation Statistics, we aimed to identify patterns in the data to answer questions such as which months or weeks have the most delayed or cancelled flights and which carriers are more likely to experience delays.}

The paper focuses mainly on how we utilized Hadoop and Spark technologies, for building the application and the Spark optimisation techniques used to improve the performance of the application. We discuss the different performance problems we faced while executing a Spark application and the measures we took to mitigate some of those issues. Specifically, we discuss the problems of data skew, spill, serialisation and shuffle. For each of these issues, we suggest techniques for optimizing Spark applications, such as data partitioning, optimizing code and choosing appropriate aggregation methods while transforming the data.

Overall, our paper demonstrates the practical use of big data technologies to extract insights from large datasets and highlights the challenges and solutions associated with processing large datasets and optimizing the execution process. The findings of our use case could be useful for travellers, airlines, airport authorities, and other stakeholders in the aviation industry to better understand the causes of flight delays and cancellations and to take appropriate measures to improve their performance.

